GROUPING AND SUMMARIZING
Summarizing the median life expentancy
```{r}
gapminder %>%
  summarise(medianLifeExp = median(lifeExp))

#Filter for 1957 then summarize the median life expectancy

gapminder %>%
  filter(year == 1957) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

```

SUMMARIZING BY YEAR
```{r}

gapminder %>%
  group_by(year) %>% #a cada año agrupado le aplica las funciones a continuación 
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap))

```

SUMMARIZING BY CONTINENT
```{r}

gapminder %>%
  filter(year == 1957) %>%
  group_by(continent) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap)) %>%
  arrange(desc(maxGdpPercap))

```

SUMMARIZING BY CONTINENT AND YEAR
```{r}
gapminder %>%
  group_by(continent, year) %>%
  summarise(medianLifeExp = median(lifeExp), maxGdpPercap = max(gdpPercap)) 
```
LINE PLOTS

VISUALIZING MEDIAN LIFE EXPECTANCY OVER TIME

```{r}
by_year <- gapminder %>%
  group_by(year) %>%
  summarise(medianGdpPercap = median(gdpPercap))

ggplot(by_year, aes(x = year, y = medianGdpPercap)) +
  geom_line()

```

```{r}
by_year_continent <- gapminder %>%
  group_by(year, continent) %>%
  summarise(medianGdpPercap = median(gdpPercap))

ggplot(by_year_continent, aes(x = year, y = medianGdpPercap, color = continent)) +
  geom_line()

```

BAR PLOTS

VISUALIZING MEDIAN GDP PER CAPITA BY CONTINENT

```{r}

by_continent <- gapminder %>%
  filter(year == 1952) %>%
  group_by(continent) %>%
  summarise(medianGdpPercap = median(gdpPercap))

  
ggplot(by_continent, aes(x=continent, y = medianGdpPercap)) +
  geom_col()


```


```{r}

oceania_1952 <- gapminder %>%
  filter(continent == "Oceania", year == 1952)

ggplot(oceania_1952, aes(x=country, y = gdpPercap))+
  geom_col()


```

HISTOGRAM

VISUALIZING POPULATION

```{r}
library(gapminder)
library(ggplot2)
library(dplyr)

gapminder_1952 <- gapminder %>%
  filter(year == 1952) %>%
  mutate(pop_by_mil = pop/1000000)

ggplot(gapminder_1952, aes(x=pop_by_mil)) +
  geom_histogram(bins = 50)

```

VISUALIZING POPULATIN WITH x = axis on a log scale

```{r}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x=pop)) +
  geom_histogram(bins = 50) +
  scale_x_log10()
         
         
```

BOXPLOTS

Comparing GDP per capita across continents

```{r}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() + 
  scale_y_log10()

```

ADDING A TITLE TO YOUR GRAPH
```{r}
gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x=continent, y=gdpPercap)) +
  geom_boxplot() + 
  scale_y_log10() +
  ggtitle("Comparing GDP Per capita across continents")
```


DATA MANIPULATION WITH DPLYR
```{r}

library(tidyverse)
counties <- read.csv('counties.csv')
glimpse(counties)


```

```{r}
counties %>%
  select(state, county, population, men, women)

counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected %>%
  arrange(desc(public_work))

counties_selected <- counties %>%
  select(state, county, population) 

counties_selected %>%
  filter(population > 1000000) 


counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)

counties_selected %>%
  filter(state == "Texas", population > 1000000) %>%
  arrange(desc(private_work))
```


```{r}
counties %>%
  select(state, county, population, men, women) %>%
  mutate(proportion_men = men/population) %>%
  filter(population > 10000) %>%
  arrange(desc(proportion_men))
```

THE VERB COUNT

```{r}

selected_counties <- counties %>%
  select(state, county, population, citizens)

counties_selected <- counties %>%
  count(region, sort = TRUE) %>%
  arrange(desc(n))

```

COUNTING CITIZIENS BY STATE

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, citizens)


counties_selected %>%
  count(state, wt = citizens) %>% #agrupa por estado, pero el conteo es por los valores de citiziens
  arrange(desc(n))
```

MUTATING AND COUNTING

```{r}
counties_selected <- counties %>%
  select(county, region, state, population, walk)

counties_selected %>%
  mutate(population_walk = population*walk/100) %>%
  count(state, wt = population_walk, sort = TRUE)
```

SUMMARIZING

```{r}

counties_selected <- counties %>%
  select(county, population, income, unemployment)

counties_selected %>%
  summarise(min_population = min(population), max_unemployment = max(unemployment), averange_income = mean(income))

```

SUMMARIZING BY STATE

```{r}
counties_selected <- counties %>%
  select(state, county, population, land_area)

counties_selected %>%
  group_by(state) %>%
  summarise(total_area = sum(land_area), total_population = sum(population)) %>%
  mutate(density = total_population/total_area) %>%
  arrange(desc(density)) 

counties_selected <- counties %>%
  select(state, region, county, population)

counties_selected %>%
  group_by(region, state) %>%
  summarise(total_population = sum(population)) %>%
  summarise(average_pop = mean(total_population), median_pop = median(total_population))

```

TOP_n VERB

```{r}

counties_selected <- counties %>%
  select(state, region, county, metro, population, walk)

counties_selected %>%
  group_by(region) %>%
  top_n(1, walk) #devuelve la mayor cantidad de porcentaje de ciudadanos caminando

counties_selected <- counties %>%
  select(state, region, county, income, population)

counties_selected %>%
  group_by(region, state) %>%
  summarise(average_income = mean(income)) %>%
  top_n(1, average_income)

```

```{r}

counties_selected <- counties %>%
  select(state, metro, population)


counties_selected %>%
  group_by(state, metro) %>%
  summarise(total_pop = sum(population)) %>%
  top_n(1, total_pop) %>%
  ungroup() %>% #quita la agrupación y convierte a un data set
  count(metro)
```

SELECTING COLUMNS

```{r}
counties %>%
  select(state, county, population, professional : production) %>%
  arrange(desc(service))

counties %>%
  select(state, county, population, ends_with("work")) %>%
  filter(public_work >= 50)

counties %>%
  select(state, county, population, starts_with("income"))

counties %>%
  select(state, county, population, contains("at"))
```

RENAMING A COLUMN AFTER COUNT

```{r}
counties %>%
  count(state) %>%
  rename(num_conties = n)


counties %>%
  select(state, county, poverty_rate = poverty)
```

USING TRANSMUTE
```{r}
counties %>%
  transmute(state, county, population, density = population/land_area) %>%
  filter(population >= 1000000) %>%
  arrange(desc(density))

```

CHOOSING AMONG THE FOUR VERBS
```{r}
#Change the name of the unemployment column

counties %>%
  rename(unemployment_rate = unemployment) 

#Keep the state and county columns, and the columns containing poverty

counties %>%
  select(state, county, contains("poverty")) #no se puede renombrar

#Calculate the fraction_women column without dropping the other column

counties %>%
  mutate(fracion_women = women/population) #no se puede filtrar

#Keep only the state, county and employment_rate columns

counties %>%
  transmute(state, county, employment_rate = employed/population) #permite filtrar y asignar variables
```

