INTRODUCTION TO THE TIDYVERSE

```{r}
library("gapminder")
library("dplyr")

gapminder
```

FILTERING FOR ONE YEAR
```{r}

gapminder %>%
  filter(year == 1957)

```

FILTERING FOR ONE COUNTRY AND ONE YEAR
```{r}
gapminder %>%
  filter(year == 1957, country == "China")
```

ARRANGING OBSERVARTIONS BY LIFE EXPECTANCY
```{r}

gapminder %>%
  arrange(lifeExp)

# Sort in descending order of lifeExp

gapminder %>%
  arrange(desc(lifeExp))
```
FILTERING AND ARRANGING
```{r}

gapminder %>%
  filter(year == 1957) %>%
  arrange(desc(pop))

```

USING MUTATE TO CHANGE OR CREATE A COLUMN

```{r}
#Use mutate to change lifeExp to be in months

gapminder %>%
  mutate(lifeExp = 12*lifeExp)

#Use mutate to create a new variable lifeExpMonths
gapminder %>%
  mutate(lifeExpMonths = 12*lifeExp)
```

COMBINING

```{r}
gapminder %>%
  filter(year == 2007) %>%
  mutate(lifeExpMOnths = 12*lifeExp) %>%
  arrange(desc(lifeExpMOnths))

```

VISUALIZING WITH ggplot2
```{r}

library(ggplot2)

gapminder_1952 <- gapminder %>%
  filter(year == 1952)

ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) + 
  geom_point()

```
```{r}
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) + 
  geom_point()
```
PUTTING THE x-Axis ON A LOG SCALE
```{r}
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() +
  scale_x_log10()
```

PUTTING THE x-and and y-axes on a log scale
```{r}
ggplot(gapminder_1952, aes(x=pop, y=gdpPercap)) +
  geom_point() +
  scale_x_log10() + 
  scale_y_log10()

```
ADDING COLOR TO A SCATTER PLOT

```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color = continent)) +
  geom_point() +
  scale_x_log10()
```

ADDING SIZE AND COLOR TO A PLOT
```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp, color = continent, size = gdpPercap)) +
  geom_point() +
  scale_x_log10()
```
FACETING
Creating a subgraph for each continent

```{r}
ggplot(gapminder_1952, aes(x=pop, y=lifeExp)) +
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ continent)

```
FACETING BY YEAR
```{r}
ggplot(gapminder, aes(x=gdpPercap, y=lifeExp, color = continent, size = pop)) +
  geom_point() +
  scale_x_log10() +
  facet_wrap(~ year)
```



